### 2014-05-01 paper.tex
Several new areas.

Advances in mathematical modelling for the control of robotic systems have also made significant advances in recent years.  While the theory is well developed1 (TODO: ???), limits in computational speed and power have kept the theory far short of practice.  One of the standard approaches for solving  the kind of optimization problems that arise in motion planning for a robot is the method of Dynamic Programming, originally developed by Richard Bellman. (TODO: ???) While the theory is powerful, the approach suffers from the curse of dimensionality \cite{random_sampling_dynamic_programming} and computational approaches fail for any realistically high-dimensional models.  The Atlas robot is very high dimensional, 28 independent hydraulic actuated joints. In addition, computational approaches to optimization can be very sensitive to model and sensor error.  There is a large literature in the area of robust dynamic programming (TODO: ???) which focuses reducing sensitivity to error.

@@

\subsubsection{Walk}

\begin{wrapfigure}{r}{0.3\textwidth}
  \begin{center}
    \includegraphics[scale=0.5]{images/step_gui.png}
  \end{center}
  \caption{The planned steps for the robot walking forward and to the left}
\end{wrapfigure}

@@

The robot is defined as tree/skeleton starting from the pelvis, defining the weight and distances between each joint. When we generate Jacobians to move the end effector of each limbs, we also give it the ability to move the pelvis on a global coordinate frame. The hand jacobians has no direct effect on the the leg joints, but indirectly moves the imaginary point that the pelvis should follow, and the legs move with it.

@@

\subsection{Inverse Dynamics}

\subsection{Results}

\subsection{Conclusion}

@@

\section{DRC Valve}
\begin{wrapfigure}{r}{0.3\textwidth}
  \begin{center}
    \includegraphics[scale=0.2]{images/turn_valve.png}
  \end{center}
  \caption{Robot in front of valves. Both shoulders right in front of each valve.}
\end{wrapfigure}
For this task the robot must close three industrial valves during a period of 30 minutes. Those consist of a 13" ball valve, and two gate valves with turning wheels of 9" and 18". The ball valve has to be closed 90 degrees clockwise (until the lever is horizontal). The gate valves have to be rotated clockwise a full turn to be considered closed.  \cite{valvetask}

\subsection{Approach}
The first problem to achieve the task is to step towards the valve. The operator must activate the point cloud feedback on the computer, and position a marker on top of what looks to be a valve on the collection of dots. With the marker in place, it gives the operator the position of where the feet must be placed wherever the robot is going to use the left hand, or the right hand, to operate the valve.

@@

\section{Conclusion}

@@

\subsection{Looking at a Problem}
To have a better understanding on how the algorithm works, we are going to look at the  inverse pendulum swing-up example, which consists of a motor attached to the end of pendulum, and the controller aims to bring the pendulum to an upright position by applying torques based on the current state of the system.

@@

\subsection{Running through an example}

@@

\subsection{Analysing this solution}
The algorithm is far from being real time, since for the single-link pendulum case, on an 8-core CPU at 2.8GHz, takes about 4 seconds for each iteration, and about 1 hour for the 1000 iterations, but it does generate very robust policy tables.

@@

\subsection{CUDA}

@@

\subsection{Applying DDP with CUDA}
Since a new improved torques can be found in parallel with other points, we can run the Dynamic Programming Algorithm using CUDA.  We initialize the process by allocating two memory buffers in the device, and reset all of the values to 0.  We create a CUDA kernel, which takes an identification number, input memory buffer, and compares the current policy with a new random policy, and stores the result on an output buffer.

@@

\section{Future Work}

>>>>

Advances in mathematical modelling for the control of robotic systems have also made significant advances in recent years.  While the theory is well developed \cite{Deterministic_and_Stochastic}, limits in computational speed and power have kept the theory far short of practice.  One of the standard approaches for solving  the kind of optimization problems that arise in motion planning for a robot is the method of Dynamic Programming, originally developed by Richard Bellman. \cite{Deterministic_and_Stochastic} While the theory is powerful, the approach suffers from the curse of dimensionality \cite{random_sampling_dynamic_programming} and computational approaches fail for any realistically high-dimensional models.  The Atlas robot is very high dimensional, 28 independent hydraulic actuated joints. In addition, computational approaches to optimization can be very sensitive to model and sensor error.

@@

\subsubsection{Walk}
 \begin{wrapfigure}{r}{0.3\textwidth}
  \begin{center}
    \includegraphics[scale=0.4]{images/step_gui.png}
  \end{center}
  \caption{The planned steps for the robot walking forward and to the left}
\end{wrapfigure}

@@

The robot is defined as tree/skeleton starting from the pelvis, defining the weight and distances between each joint. When we generate Jacobians \cite{robot_modeling} to move the end effector of each limbs, we also give it the ability to move the pelvis on a global coordinate frame. The hand jacobians has no direct effect on the the leg joints, but indirectly moves the imaginary point that the pelvis should follow, and the legs move with it.

@@

\subsection{Inverse Dynamics}
Inverse Dynamics is responsible for generating the torques applied to each joint of the Atlas robot. The equations of motion and the constraint equations for a humanoid robot can be described as:
\vspace{-5pt}
\begin{equation} \label{eq2}
\begin{split}
M(q)\ddot q + h(q,\dot q) &= S\tau + J^T(q)F\\
J(q)\ddot q + \dot J(q,\dot q)\dot q & = \ddot x
\end{split}
\end{equation}
where $(q,\dot {q})$ is the state of the system, $M(q)$ is the inertia matrix, $h(q,\dot q)$ is the sum of gravitational, centrifugal and Coriolis forces, $S$ is a selection matrix, $\tau$ is a vector of joint torques, $F$ is a vector of contact forces, and $x$ is a vector of contact positions and orientations.

\subsection{Finding grasping positions}
On an attempt to perform the debris task more efficiently, a routine was created in which the program iterates forward several instances of the Inverse Kinematics part of the controller, and checks how well the controller is able to grab the target object, and picks the one in which the center of mass is most stable within the support polygon.

Several unexpected behaviors showed from this experiment: The controller would prefer to to leave joints at the joint limits unless it was complete necessary to move it to achieve a necessary motion, and when moving the arm away from the ground, the controller will choose to move the legs first until the legs reach a singularity.

It is unclear why the controller prefers to stay at joint limits, but might be due to the way that QuadProg++ adds special constraints while at a boundary, and does not move away from the edge cases unless necessary. Although no algorithmic solution was attempted, the current solution to not to allow the initial state when sampling goal positions to have  have any joints at the limit.

Singularities are an inconvenient truth of working with Jacobians. The current solution is to simply avoid reaching singularities on the elbows and knees, or add special cases in the code. \cite{eric_thesis}

\subsection{Results}
Overall the controller is able to perform full body motions without much work from the developer, and sometimes make almost human-like behavior. Where used, it performed well on the DRC trials, and gave little trouble for the operator.

There still exists edge cases in which the controller will act unexpectedly, such as the elbows or knees reaching a singularity, joints being stuck at a boundary, and the controller performing large body motions for motions that could be accomplished only with the arms.

@@

Members at Carnegie Mellon University suggested to use SNOPT, a software package for solving large-scale optimization problems, to help with path planning. SNOPT is able to perform what QuadProg++ does, but over large motions and account for additional dynamic constraints.

Boston Dynamics has mentioned that the robot is receiving several major modifications on the arms and legs, including adding and switching the last 3 joints of each arm to be electric, instead of hydraulic, and providing new legs. All of these changes are bound to be completed by the end of January, which provides several complications in developing software that works specifically for Atlas in his current state.

\section{DRC Valve}

For this task the robot must close three industrial valves during a period of 30 minutes. Those consist of a 13" ball valve, and two gate valves with turning wheels of 9" and 18". The ball valve has to be closed 90 degrees clockwise (until the lever is horizontal). The gate valves have to be rotated clockwise a full turn to be considered closed.  \cite{valvetask}

\subsection{Approach}
\begin{wrapfigure}{r}{0.3\textwidth}
  \begin{center}
    \includegraphics[scale=0.2]{images/turn_valve.png}
  \end{center}
  \caption{Robot in front of valves. Both shoulders right in front of each valve.}
\end{wrapfigure}
The first problem to achieve the task is to step towards the valve. The operator must activate the point cloud feedback on the computer, and position a marker on top of what looks to be a valve on the collection of dots. With the marker in place, it gives the operator the position of where the feet must be placed wherever the robot is going to use the left hand, or the right hand, to operate the valve.

@@

@@

\section{Looking at a Problem}
To have a better understanding on how the algorithm works, we are going to look at the  inverse pendulum swing-up example, which consists of a motor attached to the end of pendulum, and the controller aims to bring the pendulum to an upright position by applying torques based on the current state of the system.

@@

\section{Running through an example}

@@

\section{Analysing this solution}
The algorithm is far from being real time, since for the single-link pendulum case, on an 8-core CPU at 2.8GHz, takes about 4 seconds for each iteration, and about 1 hour for the 1000 iterations, but it does generate very robust policy tables.

@@

\section{CUDA}

@@

\section{Applying DDP with CUDA}
Since a new improved torques can be found in parallel with other points, we can run the Dynamic Programming Algorithm using CUDA.  We initialize the process by allocating two memory buffers in the device, and reset all of the values to 0.  We create a CUDA kernel, which takes an identification number, input memory buffer, and compares the current policy with a new random policy, and stores the result on an output buffer.

@@

\chapter{Future Work}
